---
title: "SanderBeerepoot"
format: html
---

```{r}
library(here)
library(vroom)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
```

```{r}
here::i_am("eval_grades.Rproj")
```

```{r}
courses <- vroom(here("courses.csv"))
```

Question 2
Display the content of the file in a nicely formatted table in your document. Courses should be
sorted in alphabetical order and column names should be written in English. This has to be
done in R for instance using knitr::kable().

```{r}
formatted_courses <-
courses |>
  rename(Course_name = course,
         identifier = course_id,
         module = module,
         `number of exams` = nb_grades)|>
  arrange(Course_name)
knitr::kable(formatted_courses)
```
```{r}
students <- 
  vroom(here("students.csv"))
```
to have group loaded as a factor and id as an integer.
Question 3
Load the student data set. Make sure the birth dates are correctly identified as Date objects by
including the class of the corresponding column in the rendered text. For instance, include a
sentence of the following form:
The birth_date column of the students data frame is of class Date.
Make also sure that the sex is identified as a factor.
```{r}
students_typed <- students |>
  mutate(
    birth_date = as.Date(birth_date, format = "%Y-%m-%d"),
    group = as.factor(group),
    sex = as.factor(sex),
    id = as.integer(id)
  )
```

```{r}
grades <- vroom(here("grades.csv"))
cols(
  id = col_double(),
  course_id = col_double(),
  grade = col_double())
```
Question 4

```{r}
ok_grades <- read_delim(here("grades.csv"),
  delim = "|"
)
cols(
  id = col_double(),
  course_id = col_double(),
  grade = col_double())
```
Question 5: Use a graphical representation to display the number of students per group.
```{r}
students_typed |>
  group_by(group)|>
  summarise(n = n())|>
  ggplot(aes(x = n)) +
  geom_bar()
```
 Question 6: Show the gender balance in each group on a single graphical representation.
```{r}
students_typed|>
  ggplot(aes(x = group, fill = sex)) +
  geom_bar(position = "fill")
```

Question 7
Show graphically the distribution of the age of the students.

```{r}
students_age <- students|>
  mutate(age_years = as.numeric(Sys.Date() - birth_date) / 365.25)
students_age |>
ggplot(aes(x = age_years)) +
  geom_density()
  

```
Question 8
Compute the median age of the students in each group and include in your rendering a table
giving the median age for each group. The median should be reported as an integer.

```{r}
students_age |>
  group_by(group)|>
  summarise(median = median(age_years))
```

